## Set working directory and (install and) load any packages

```{r setup}
knitr::opts_knit$set(root.dir = "~/network-storage/WNV/sjv-wnv/")

library(ggplot2)
library(naniar)
library(cowplot)
library(plm)
library(dplyr)
library(lmtest)
library(pglm)
library(tidyr)
library(ggforce)
```



## Data

For the sake of simplicity, we will use only the data buffered at 1500 meters.

```{r}
bird_data <- read.csv("Data/birds/bird1500Lag0.csv") %>% 
  arrange(clust, Year) %>% 
  group_by(clust, Year) %>% 
  arrange(Month, .by_group = T) %>% 
  ungroup() %>% 
  mutate(m = (Year * 12 - 24072) + Month)
  
pool_data <- read.csv("Data/pools/pool1500Lag0.csv") %>% 
  filter((Count + CountLinear + CountSpline) > 0) %>% 
  arrange(clust, Year) %>% 
  group_by(clust, Year) %>% 
  arrange(Month, .by_group = T) %>% 
  ungroup() %>% 
  mutate(m = (Year * 12 - 24072) + Month)

sentinel_data <- read.csv("Data/sentinels/sentinel1500Lag0.csv") %>% 
  arrange(clust, Year) %>% 
  group_by(clust, Year) %>% 
  arrange(Month, .by_group = T) %>% 
  ungroup() %>% 
  mutate(m = (Year * 12 - 24072) + Month)

```

## Bird data

```{r}
ggplot(bird_data) +
  geom_histogram(aes(Count))
```

### Simple Regression Model

```{r}
bird_data_pd <- pdata.frame(bird_data, index = c("clust", "m"))

# estimate simple regression model
bird_data_lm <- lm(Count ~ jrcStandingWater, data = bird_data_pd)

coeftest(bird_data_lm, vcov. = vcovHC, type = "HC1")

ggplot() + geom_density(aes(bird_data_lm$residuals))
```

Based on the simple regression model, as standing water increases, the count of infected birds increase per cluster. However, the residuals are not normally distributed, which makes sense since our target variable is discrete.

Now we try using a panel linear model using the `plm()` function.

### Regression with Fixed Effects

```{r}
# estimate the fixed effects regression with plm()

bird_fe_plm <- plm(
  Count ~ jrcStandingWater,
  data = bird_data_pd,
  index = c("clust", "m"),
  model = "within"
)

coeftest(bird_fe_plm, vcov. = vcovHC, type = "HC1")

ggplot() +
  geom_density(aes(bird_fe_plm$residuals))
```

The coefficient for `jrcStandingWater` is not statistically significant, so we move on to another method.

### Regression with Time-Fixed Effects


```{r}
bird_tefe_plm <- plm(
  Count ~ jrcStandingWater,
  data = bird_data_pd,
  index = c("clust", "m"),
  model = "within",
  effect = "twoways"
)

coeftest(bird_tefe_plm, vcov. = vcovHC, type = "HC1")

summary(bird_tefe_plm)

ggplot() +
  geom_density(aes(bird_tefe_plm$residuals))
```

The coefficient is still not statistically significant, though it is an improvement from before. 

Now, let's add some covariates.

### Several Various Models

```{r}
bird_mod1 <- pglm(
  Count ~ jrcStandingWater, 
  data = bird_data_pd, 
  family = poisson(link = 'sqrt')
)

summary(bird_mod1)

bird_mod2 <- pglm(
  Count ~ jrcStandingWater + Year + Month,
  data = bird_data_pd,
  family = poisson(link = 'sqrt')
)

summary(bird_mod2)


bird_mod3 <- plm(
  Count ~ jrcStandingWater + EVI + Precip.Mean.mm...day + pdsi + Specific.Humidity,
  data = bird_data_pd
)

summary(bird_mod3)

ggplot() + geom_density(aes(bird_mod3$residuals))

bird_mod4 <- plm(
  Count ~ jrcStandingWater + EVI + Precip.Mean.mm...day + pdsi + Specific.Humidity,
  data = bird_data_pd,
  effect = "twoways",
  model = "within"
)

summary(bird_mod4)

ggplot() + geom_density(aes(bird_mod4$residuals))

bird_mod4 <- plm(
  Count ~ jrcStandingWater + Precip.Mean.mm...day + pdsi + Specific.Humidity + z,
  data = bird_data_pd,
  index = c("clust", "m"),
  model = "within"
)


coeftest(bird_mod4)
summary(bird_mod4)

ggplot() + geom_density(aes(bird_mod4$residuals))

bird_mod5 <- plm(
  log(Count) ~ jrcStandingWater + Precip.Mean.mm...day + pdsi + Specific.Humidity + z,
  data = bird_data_pd,
  index = c("clust", "m"),
  effect = "twoways",
  model = "within"
)

summary(bird_mod5)

coeftest(bird_mod5, vcov. = vcovHC, type = "HC1")

ggplot() + geom_density(aes(bird_mod5$residuals))

bird_mod6 <- glm(
  Count ~ jrcStandingWater + Year - 1 + + NDVI + Precip.Mean.mm...day + pdsi + Specific.Humidity +
    Humidity.Max + Humidity.Min + z + spi1y, 
  data = bird_data_pd,
  family = poisson
)

summary(bird_mod6)

coeftest(bird_mod6, vcov. = vcovHC, type = "HC1")

ggplot() + geom_density(aes(bird_mod6$residuals))
```


```{r}
bird_data_factored <- bird_data %>% 
  mutate(Count = as.factor(Count)) %>% 
  group_by(clust, m, Count) %>% 
  pivot_longer(c(Month, Year, Humidity.Max:z)) %>% 
  ungroup() %>% 
  select(clust, Count, m, name, value)

# put in number of plots here
noPlots <- 1
# save all variables in a seperate vector to select in for-loop
allVars <- unique(bird_data_factored$name)
noVars <- length(allVars)

# indices for plotting variables
plotSequence <- c(seq(0, noVars-1, by = noPlots), noVars)

# pdf("plotpath.pdf") # uncomment to save the resulting plots in a pdf file
# loop over the variables to plot
for(ii in 2:length(plotSequence)){
  # select start and end of variables to plot
  start <- plotSequence[ii-1] + 1
  end <- plotSequence[ii]

  # subset the variables and save new temporary data.frame
  tmp <- subset(bird_data_factored, name %in% allVars[start:end])
  cat(unique(tmp$name), "\n")

  # generate plot
  p <- ggplot(tmp)+
    geom_boxplot(aes(value)) +
    facet_wrap( ~ Count , ncol = 5, scales = "fixed") +
    xlab(tmp$name)
  print(p)
}

```

